---
title: "Of Fights and Lights: Conflict and Economic Activity in Thailand"
author:
  - "Gustav Pirich (h11910449)"
date: "May 4, 2024"
output:
  pdf_document:
    toc: true
    includes:
      in_header: !expr file.path("~/Desktop/GITHUB/spatial_econ/helper/wraper_code.tex")
bibliography: references.bib
nocite: '@*'
header-includes:
  - \usepackage{tcolorbox}
  - \usepackage[default]{lato}
  - \usepackage{rotating}
  - \usepackage{dcolumn}
  - \usepackage{booktabs}
  - \usepackage{float}
papersize: a4
geometry: margin=2cm
urlcolor: DarkOrchid!65!black
---


```{r, setup, echo = FALSE, warning=FALSE, results = 'hide'}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

pacman::p_load(spatialreg, showtext, magritr, bsreg, spdep, patchwork, gridExtra, fixest, splm, stringi, stringr, stringdist, haven, sf, dplyr, fuzzyjoin, comparator, digest, zoomerjoin, ggplot2, tidyr, ggthemes, viridis, rmapshaper, fixest, conleyreg, plm, stargazer, bsreg, igraph, generics, knitr, kableExtra, formatR,readxl, haven, flextable, broom, units, sysfonts, showtextdb, tmap, viridis)
```

\vspace{2em}

\begin{tcolorbox}
\centering \itshape The code that was used in compiling the assignment is available on GitHub at \url{https://github.com/gustavpirich/thai_conflict}.
\end{tcolorbox}


\newpage 

To investigate the connection between conflicts and nighttime lights in Thailand, I use nighttime light data at the province level. Thailand has a total number of 928 provinces. I use DSMP nighttime light data for the period from 1992 up until 2011. For the period from 2012 to 2021, I use VIIRS nighttime lights data. Both datasets are sourced from @aiddata_2021. I present three variables; the mean intensity of nighttime lights in a the first and last year of the sample, the growth rate of nighttime lights over a period, and the number of conflict events.


# DSMP Nighttime Lights from 1992-2011

The panel on the left displays nighttime light intensities in Thailand. Economic activity is strongly concentrated in Bangkok, the economic and political center of the country. The second panel displays the relative change of nighttime lights from 1992 to 2011. The panel on the right presents the number of reported conflict events over the time period. \footnote{Note that DSMP data suffers from topcoding, which might lead to an understatement of growth rates in urban areas such as Bangkok}

From the third map, it can be seen that conflict events are heavily concentrated in the Southern provinces.  The conflict events refer to the long-standing ethnic and religious tensions and insurgencies by Islamist groups. The first panel demonstrates that the Southern part of the country is likely to be poor and agrarian as proxied by nighttime lights. However, the provinces affected by conflict events seem to have grown at an even faster pace than  the rest of the country. 

```{r, echo = FALSE, include=FALSE}

# Read shapefiles for national and first administrative level geometries
thai_0 <- read_sf("gadm41_THA_shp/gadm41_THA_0.shp")
thai_1 <- read_sf("gadm41_THA_shp/gadm41_THA_1.shp")
thai_2 <- read_sf("gadm41_THA_shp/gadm41_THA_2.shp")

thai_0_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM0-all/geoBoundaries-THA-ADM0.shp")
thai_1_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM1-all/geoBoundaries-THA-ADM1.shp")
thai_2_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM2-all/geoBoundaries-THA-ADM2.shp")

# Read the conflict data
conflict <- readRDS("ged241.rds")

# Filter thai conflict data and convert to spatial dataframe
thai_conflict <- conflict %>%
  filter(country == "Thailand") %>%
  st_as_sf(coords = c("longitude", "latitude"))

# Set coordinate reference system for spatial data
thai_conflict <- st_set_crs(thai_conflict, 4326)

# Define a sequence of years for analysis
year <- 1992:2020

# Create a dataframe with all combinations of GID and year
expanded_df <- expand.grid(thai_1$GID_1, year)
names(expanded_df) <- c("GID_1", "year")

# Spatial join to match conflicts to provinces using geometry
joined_data <- st_join(thai_conflict, thai_1 %>% select(GID_1, geometry), join = st_within)

# Create a conflict event panel data
conflict_panel <- expanded_df %>%
  left_join(joined_data, by = c("GID_1", "year"))

# Summarize number of conflict events per province per year
simplified_data <- conflict_panel %>%
  group_by(GID_1, year) %>%
  summarise(number_of_events = sum(!is.na(conflict_name)), .groups = "drop") %>%
  left_join(thai_1 %>% select(GID_1, geometry))

# Simplify the geometries to improve performance of plotting
thai_1_simplified <- thai_1 %>%
  st_simplify(dTolerance = 1000) %>%
  select(GID_1, geometry)

```


```{r, echo = FALSE}
thai_2_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM2-all/geoBoundaries-THA-ADM2.shp")

adm2_dsmp <- read.csv("adm2_dsmp.csv")

# Transform to long format
adm2_dsmp_long <- adm2_dsmp %>%
  pivot_longer(
    cols = starts_with("v4composites_calibrated_201709"), # Update prefix to match new dataset
    names_to = "year", # name of the new key column
    values_to = "mean_intensity" # name of the new value column
  )  %>% 
  mutate(year = as.numeric(gsub("v4composites_calibrated_201709\\.|\\.mean", "", year))) 


# Extract the year from the column name
adm2_dsmp_long_prep <- adm2_dsmp_long %>%
  group_by(shapeID, shapeName) %>%
  mutate(mean_intensity = mean_intensity + 1) %>%
  filter(year %in% c(1992, 2011)) %>% # Adjust the years based on the new dataset
  summarise(
    growthRate = (mean_intensity[year == 2011] - mean_intensity[year == 1992]) / mean_intensity[year == 1992]
  ) 

# downsampling the adminstrative units of thai_2_adm
thai_2_adm_simplify <- thai_2_adm %>% 
  st_simplify(dTolerance = 500)

geo_adm2_growth <- left_join(adm2_dsmp_long_prep, thai_2_adm_simplify, by = c("shapeName"))

geo_adm2_growth <- geo_adm2_growth %>%
  st_as_sf 

######### Data Preparation for Lights 


adm2_dsmp_long_2011 <- adm2_dsmp_long %>%
  filter(year == 1992) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string

# downsampling the adminstrative units of thai_2_adm
thai_2_adm_simplify <- thai_2_adm %>% 
  st_simplify(dTolerance = 500)

geo_adm2_wealth <- cbind(adm2_dsmp_long_2011, thai_2_adm_simplify)

geo_adm2_wealth_sf <- geo_adm2_wealth %>%
  st_as_sf 
####################################  

# Read the conflict data
conflict <- readRDS("ged241.rds")

# Filter thai conflict data and convert to spatial dataframe
thai_conflict <- conflict %>%
  filter(country == "Thailand") %>%
  st_as_sf(coords = c("longitude", "latitude"))

# Set coordinate reference system for spatial data
thai_conflict <- st_set_crs(thai_conflict, 4326)

# filtering for Thai conflict events form 1992 to 2011
thai_conflict_viirs <- thai_conflict %>%
  filter(year %in% c(1992:2011))

# Perform a spatial join to count conflict events within each province
conflict_counts <- thai_conflict_viirs %>%
  st_join(thai_2_adm_simplify, join = st_within) %>%
  group_by(shapeName) %>%  # Adjust this to your province name field
  summarise(conflict_count = n())

# Merge conflict counts back to the provinces dataset
thai_provinces <- thai_2_adm_simplify %>%
  st_join(conflict_counts) %>%
  mutate(conflict_count = ifelse(is.na(conflict_count), 0, conflict_count))
###########



wealth <- tm_shape(geo_adm2_wealth_sf) +
  tm_fill(
    "mean_intensity",
    palette = "magma", 
    #style = "bclust",
    breaks = c(0, 10, 20, 30, 40, 50, 63),
    border.col = "black",
    border.alpha = 0.2,
    legend.frame = FALSE, 
    title.fontface = 2
  ) +  
  tm_borders(col = "#333333", lwd = 0.11) +
  tm_layout(
    frame = FALSE, 
    main.title = "Nighttime Light Intensity in 2000",
    main.title.size = 0.6, 
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    #legend.outside=TRUE,
    legend.position = c("right", "bottom"), 
    legend.frame = FALSE, 
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )  


conflicts_provincial <- tm_shape(thai_provinces) +
  tm_fill(
    "conflict_count",
    palette = "magma", 
    border.col = "black",
    border.alpha = 0.2,
    legend.frame = FALSE, 
    title.fontface = 2, 
    breaks = c(0,5,10,15,20,25,30,50)
  ) +  
  tm_borders(col = "#333333", lwd = 0.11) +
  tm_layout(
    frame = FALSE, 
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    legend.frame = FALSE, 
    #legend.outside=TRUE,
    legend.position = c("right", "bottom"),     
    main.title = "Conflict Events from 1992-2011",
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    main.title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )  

growth <- tm_shape(geo_adm2_growth) +
  tm_polygons(
    "growthRate",
    palette = "magma", 
    #style = "sd",
    breaks = c(-2, 0, 2, 4, 6, 8, 10, Inf), # Custom breaks
    #labels <- c("< -4", "-4 to -2", "-2 to 0", "0 to 2", "2 to 4", "4 to 6", "6 to 8", "> 8"), 
    border.col = "black",
    border.alpha = 0.2, 
    title.fontface = 2
  ) +
  tm_layout(
    main.title = "NTL Growth from 1992 to 2011",
    frame = FALSE, 
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    #legend.outside=TRUE,
    legend.position = c("right", "bottom"), 
    legend.frame = FALSE, 
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    main.title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )


tm_dsmp <- tmap_arrange(wealth, growth, conflicts_provincial, ncol = 3)
tm_dsmp
```



# VIIRS Nighttime Lights 2012-2021

For the second panel I use province-level VIIRS nighttime lights, which are of signficantly higher quality than the DSMP product @Gibson_2021.  

The third panel again displays the conflicts raging in the Southern provinces of Thailand. The first panel shows that these provinces can still be considered to be poor. The second panel displays slower growth rates for the period from 2012 to 2021. Thus, there seems to be a change in the relationship between conflict events and nighttime light growth as compared to the period from 1992 to 2011.  

```{r, echo = FALSE}
######### Data Preparation for Lights 

# Reading in VIIS data
adm2_viirs <- read.csv("adm2_viirs.csv")

adm2_viirs_long <- adm2_viirs %>%
  pivot_longer(
    cols = starts_with("viirs_ntl_annual_v21_avg_masked"), # specify the prefix common to all year columns
    names_to = "year", # name of the new key column
    values_to = "mean_intensity" # name of the new value column
  ) %>%
  mutate(year = as.numeric(gsub("viirs_ntl_annual_v21_avg_masked\\.|\\.mean", "", year)))

# Filter VIIRS data for 2014 to plot wealth
adm2_viirs_2021 <- adm2_viirs_long %>%
  filter(year == 2021) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string

# downsampling the adminstrative units of thai_2_adm
thai_2_adm_simplify <- thai_2_adm %>% 
  st_simplify(dTolerance = 500)

geo_adm2_wealth <- cbind(adm2_viirs_2021, thai_2_adm_simplify)

geo_adm2_wealth_sf <- geo_adm2_wealth %>%
  st_as_sf 
####################################  


######### Data Preparation for Conflict 
thai_2_adm_simplified <- thai_2_adm %>% 
  st_simplify(dTolerance = 500)

# Read the conflict data
conflict <- readRDS("ged241.rds")

# Filter thai conflict data and convert to spatial dataframe
thai_conflict <- conflict %>%
  filter(country == "Thailand") %>%
  st_as_sf(coords = c("longitude", "latitude"))

# Set coordinate reference system for spatial data
thai_conflict <- st_set_crs(thai_conflict, 4326)

# filtering for Thai conflict events form 2012 to 2021
thai_conflict_viirs <- thai_conflict %>%
  filter(year >= 2012)

# Perform a spatial join to count conflict events within each province
conflict_counts <- thai_conflict_viirs %>%
  st_join(thai_2_adm_simplified, join = st_within) %>%
  group_by(shapeName) %>%  # Adjust this to your province name field
  summarise(conflict_count = n())

# Merge conflict counts back to the provinces dataset
thai_provinces <- thai_2_adm_simplified %>%
  st_join(conflict_counts) %>%
  mutate(conflict_count = ifelse(is.na(conflict_count), 0, conflict_count))
###########





adm2_viirs <- read.csv("adm2_viirs.csv")

adm2_viirs_long <- adm2_viirs %>%
  pivot_longer(
    cols = starts_with("viirs_ntl_annual_v21_avg_masked"), # specify the prefix common to all year columns
    names_to = "year", # name of the new key column
    values_to = "mean_intensity" # name of the new value column
  ) %>%
  mutate(year = as.numeric(gsub("viirs_ntl_annual_v21_avg_masked\\.|\\.mean", "", year)))

# Extract the year from the column name
adm2_viirs_long_prep <- adm2_viirs_long  %>%
  group_by(shapeID) %>%
  filter(year %in% c(2012, 2021)) %>%
  summarise(
    growthRate = (mean_intensity[year == 2021] - mean_intensity[year == 2012]) / mean_intensity[year == 2012]
  ) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string


# downsampling the adminstrative units of thai_2_adm
thai_2_adm_simplify <- thai_2_adm %>% 
  st_simplify(dTolerance = 1000)

geo_adm2_growth <- cbind(adm2_viirs_long_prep, thai_2_adm_simplify)

geo_adm2_growth <- geo_adm2_growth %>%
  st_as_sf 


wealth <- tm_shape(geo_adm2_wealth_sf) +
  tm_fill(
    "mean_intensity",
    palette = "magma", 
    breaks = c(0, 2, 4, 6, 8, 10, Inf),
    border.col = "black",
    border.alpha = 0.2,
    legend.frame = FALSE, 
    title.fontface = 2
  ) +  
  tm_borders(col = "#333333", lwd = 0.11) +
  tm_layout(
    frame = FALSE, 
    main.title = "Nighttime Light Intensity in 2021",
    main.title.size = 0.6, 
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    legend.position = c("right", "bottom"), 
    #legend.outside=TRUE,
    legend.frame = FALSE, 
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )  
  

conflicts_provincial <- tm_shape(thai_provinces) +
  tm_fill(
    "conflict_count",
    palette = "magma", 
    border.col = "black",
    border.alpha = 0.2,
    legend.frame = FALSE, 
    title.fontface = 2, 
    breaks = c(0,5,10,15,20,25,30,50)
  ) +  
  tm_borders(col = "#333333", lwd = 0.11) +
  tm_layout(
    frame = FALSE, 
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    legend.position = c("right", "bottom"), 
    #legend.outside=TRUE,
    legend.frame = FALSE, 
    main.title = "Conflict Events from 2012-2021",
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    main.title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )  

growth <- tm_shape(geo_adm2_growth) +
  tm_polygons(
    "growthRate",
    palette = "magma", 
    breaks = c(-2, 0, 2, 4, 6, 8, 10, 12, Inf), # Custom breaks
    #labels <- c("< -4", "-4 to -2", "-2 to 0", "0 to 2", "2 to 4", "4 to 6", "6 to 8", "> 8"), 
    border.col = "black",
    border.alpha = 0.2, 
    title.fontface = 2
  ) +
  tm_layout(
    main.title = "NTL Growth from 2012 to 2021",
    frame = FALSE, 
    #title = "Percentage Points",
    #fontfamily = "Arial",
    outer.bg.color = "transparent", 
    bg.color = "transparent",
    legend.position = c("right", "bottom"), 
    #legend.outside=TRUE,
    legend.frame = FALSE, 
    #scale = 0.7,  # Smaller title size
    main.title.position = c("left", "top"), 
    main.title.size = 0.6,  # Adjust this value to reduce legend title size
    legend.title.size = 0.4,  # Adjust this value to reduce legend title size
    legend.text.size = 0.4  # Adjust this value to reduce legend text size
  )


tm <- tmap_arrange(wealth, growth, conflicts_provincial, ncol = 3)
tm
```


```{r, echo = FALSE}
thai_2_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM2-all/geoBoundaries-THA-ADM2.shp")

adm2_dsmp <- read.csv("adm2_dsmp.csv")

# Transform to long format
adm2_dsmp_long <- adm2_dsmp %>%
  pivot_longer(
    cols = starts_with("v4composites_calibrated_201709"), # Update prefix to match new dataset
    names_to = "year", # name of the new key column
    values_to = "mean_intensity" # name of the new value column
  )  %>% 
  mutate(year = as.numeric(gsub("v4composites_calibrated_201709\\.|\\.mean", "", year))) 


# Extract the year from the column name
adm2_dsmp_long_prep <- adm2_dsmp_long %>%
  group_by(shapeID, shapeName) %>%
  mutate(mean_intensity = mean_intensity + 1) %>%
  filter(year %in% c(1992, 2011)) %>% # Adjust the years based on the new dataset
  summarise(
    growthRate = (mean_intensity[year == 2011] - mean_intensity[year == 1992]) / mean_intensity[year == 1992]
  ) 

######### Data Preparation for Lights 

adm2_dsmp_long_2011 <- adm2_dsmp_long %>%
  filter(year == 2011) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string


###### conflict data

# Filter thai conflict data and convert to spatial dataframe
thai_conflict <- conflict %>%
  filter(country == "Thailand") %>%
  st_as_sf(coords = c("longitude", "latitude"))

# Set coordinate reference system for spatial data
thai_conflict <- st_set_crs(thai_conflict, 4326)

# filtering for Thai conflict events form 2012 to 2021
thai_conflict_dsmp <- thai_conflict %>%
  filter(year <= 2011)

# Perform a spatial join to count conflict events within each province
conflict_counts <- thai_conflict_dsmp %>%
  st_join(thai_2_adm_simplified, join = st_within) %>%
  group_by(shapeName) %>%  # Adjust this to your province name field
  summarise(conflict_count = n())

# Merge conflict counts back to the provinces dataset
thai_provinces <- thai_2_adm_simplified %>%
  st_join(conflict_counts) %>%
  mutate(conflict_count = ifelse(is.na(conflict_count), 0, conflict_count)) %>%
  rename(shapeName = shapeName.x)

# Identify unique geometries
is_unique <- !duplicated(thai_provinces$geometry)

# Filter the dataset by unique geometries
thai_provinces_unique <- thai_provinces[is_unique, ]


# merging growth data
adm2_growth <- adm2_dsmp_long_prep %>%
  arrange(shapeName)

thai_wealth_conflict <- cbind(thai_provinces_unique, adm2_dsmp_long_2011)
thai_wealth_conflict_growth <- cbind(thai_wealth_conflict, adm2_growth)

########################## Data Preparation finished for wealth ########################## 
thai_wealth_conflict_growth <- thai_wealth_conflict_growth %>% 
  select(shapeName, growthRate, mean_intensity, conflict_count) %>%
  st_drop_geometry()





# Define function to calculate summary statistics
summary_stats <- function(data) {
  data %>%
    summarise(
      n = n(),
      "Mean Growth" = mean(growthRate, na.rm = TRUE),
      "SD Growth" = sd(growthRate, na.rm = TRUE),
      "Mean Intensity"  = mean(mean_intensity, na.rm = TRUE),
      "SD Intensity" = sd(mean_intensity, na.rm = TRUE),
      "Mean Conflict Count" = mean(conflict_count, na.rm = TRUE))
}

# Calculate summary statistics for the entire dataset
summary_all <- summary_stats(thai_wealth_conflict_growth)

# Split the data by conflict event presence
df_conflict <- thai_wealth_conflict_growth %>% filter(conflict_count > 0)
df_no_conflict <- thai_wealth_conflict_growth %>% filter(conflict_count == 0)

# Calculate summary statistics for both groups
summary_conflict <- summary_stats(df_conflict)
summary_no_conflict <- summary_stats(df_no_conflict)

# Combine summaries into one data frame
summary_df <- rbind(
  data.frame(Group = "All Data", summary_all),
  data.frame(Group = "With Conflict", summary_conflict),
  data.frame(Group = "Without Conflict", summary_no_conflict)
)

colnames(summary_df) <- c("Group", 
                          "n", 
                          "Mean Growth", 
                          "SD Growth", 
                          "Mean Intensity", 
                          "SD Intensity", 
                          "Mean Conflict Events")

kable(summary_df, caption = "DSMP 1992-2011", digits = 2) %>%
  kable_styling(font_size = 10) %>%  # Set font size to 10
  row_spec(0, bold = TRUE) %>%       # Make the header row bold
  kable_styling(latex_options = "striped", full_width = F) %>%  # Optional: Add stripes for better readability
  kable_styling(position = "center") %>%  # Center align the table
  scroll_box(width = "80%", height = "300px")  # Add scroll box if the table is too wide or tall




```

\newpage

```{r, echo = FALSE, fig.pos = "H"}
thai_2_adm <- read_sf("/Users/gustavpirich/Desktop/GITHUB/thai_conflict/geoBoundaries-THA-ADM2-all/geoBoundaries-THA-ADM2.shp")

# downsampling the adminstrative units of thai_2_adm
thai_2_adm_simplify <- thai_2_adm %>% 
  st_simplify(dTolerance = 500)


adm2_viirs <- read.csv("adm2_viirs.csv")

adm2_viirs_long <- adm2_viirs %>%
  pivot_longer(
    cols = starts_with("viirs_ntl_annual_v21_avg_masked"), # specify the prefix common to all year columns
    names_to = "year", # name of the new key column
    values_to = "mean_intensity" # name of the new value column
  ) %>%
  mutate(year = as.numeric(gsub("viirs_ntl_annual_v21_avg_masked\\.|\\.mean", "", year)))

# Filter VIIRS data for 2014 to plot wealth
adm2_viirs_2021 <- adm2_viirs_long %>%
  filter(year == 2021) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string


# Extract the year from the column name
adm2_viirs_long_prep <- adm2_viirs_long  %>%
  group_by(shapeID) %>%
  filter(year %in% c(2012, 2021)) %>%
  summarise(
    growthRate = (mean_intensity[year == 2021] - mean_intensity[year == 2012]) / mean_intensity[year == 2012]
  ) %>%
  mutate(shapeID = str_extract(shapeID, "B\\d+")) %>%
  mutate(shapeID = str_sub(shapeID, 2)) # Remove the 'B' from the extracted string





# Read the conflict data
conflict <- readRDS("ged241.rds")

# Filter thai conflict data and convert to spatial dataframe
thai_conflict <- conflict %>%
  filter(country == "Thailand") %>%
  st_as_sf(coords = c("longitude", "latitude"))

# Set coordinate reference system for spatial data
thai_conflict <- st_set_crs(thai_conflict, 4326)

# filtering for Thai conflict events form 2012 to 2021
thai_conflict_viirs <- thai_conflict %>%
  filter(year >= 2012)

# Perform a spatial join to count conflict events within each province
conflict_counts <- thai_conflict_viirs %>%
  st_join(thai_2_adm_simplify, join = st_within) %>%
  group_by(shapeName) %>%  # Adjust this to your province name field
  summarise(conflict_count = n())

# Merge conflict counts back to the provinces dataset
thai_provinces <- thai_2_adm_simplify %>%
  st_join(conflict_counts) %>%
  mutate(conflict_count = ifelse(is.na(conflict_count), 0, conflict_count))
###########


# Identify unique geometries
is_unique <- !duplicated(thai_provinces$geometry)

# Filter the dataset by unique geometries
thai_provinces_unique <- thai_provinces[is_unique, ]



############## Preparing summary statistics dataset ############## 
growth_wealth <- left_join(adm2_viirs_2021, adm2_viirs_long_prep)

growth_wealth_conflict <-  cbind(thai_provinces_unique, growth_wealth)

growth_wealth_conflict_final <- growth_wealth_conflict %>%
  select(shapeName, mean_intensity, growthRate, conflict_count) %>%
  st_drop_geometry()
###################################################################

# Define function to calculate summary statistics
summary_stats <- function(data) {
  data %>%
    summarise(
      n = n(),
      "Mean Growth" = mean(growthRate, na.rm = TRUE),
      "SD Growth" = sd(growthRate, na.rm = TRUE),
      "Mean Intensity"  = mean(mean_intensity, na.rm = TRUE),
      "SD Intensity" = sd(mean_intensity, na.rm = TRUE),
      "Mean Conflict Count" = mean(conflict_count, na.rm = TRUE))
}

# Calculate summary statistics for the entire dataset
summary_all <- summary_stats(growth_wealth_conflict_final)

# Split the data by conflict event presence
df_conflict <- growth_wealth_conflict_final %>% filter(conflict_count > 0)
df_no_conflict <- growth_wealth_conflict_final %>% filter(conflict_count == 0)

# Calculate summary statistics for both groups
summary_conflict <- summary_stats(df_conflict)
summary_no_conflict <- summary_stats(df_no_conflict)

# Combine summaries into one data frame
summary_df <- rbind(
  data.frame(Group = "All Data", summary_all),
  data.frame(Group = "With Conflict", summary_conflict),
  data.frame(Group = "Without Conflict", summary_no_conflict)
)


colnames(summary_df) <- c("Group", 
                          "n", 
                          "Mean Growth", 
                          "SD Growth", 
                          "Mean Intensity", 
                          "SD Intensity", 
                          "Mean Conflict Events")

# Create the kable table with reduced size
kable(summary_df, caption = "VIIRS 2012-2021", digits = 2) %>%
  kable_styling(font_size = 10) %>%  # Set font size to 10
  row_spec(0, bold = TRUE) %>%       # Make the header row bold
  kable_styling(latex_options = "striped", full_width = F) %>%  # Optional: Add stripes for better readability
  kable_styling(position = "center") %>%  # Center align the table
  scroll_box(width = "80%", height = "300px")  # Add scroll box if the table is too wide or tall

```


The summary statistics corroborate the intuition gained from the maps. NTL emission in provinces were conflict events were reported grew by 261 percent from 1992 to 2011. NTL emission from peaceful provinces grew by 179 percent over the whole period.

This relationship changed from 2012 up until 2021. Provinces were conflicts were reported increased nighttime light emissions on average by 154%, whereas in peaceful provinces nighttime lights increased by 199%.

The relationship in this context cannot be interpreted as causal. To identify a causal effect of conflict on economic activity, we would need to account for various endogeneity concerns like omitted variable bias and reverse causality.  

To identify the causal effect of conflict on economic activity one could try to estimate the impact in a panel regression using province and time fixed effects, while also controlling for other time-varying factors. The timing of the onset of conflict events might serve as the 'plausibly exogenous' treatment. One could then estimate an event study for the local impact of insurgencies on economic activity. The validity of this identification scheme hinges on the parallel trends assumption. One would need to assume that provinces (or localities) in the absence of conflict would have grown at the same rate. The assumption of this staggered diff-in-diff identification strategy will be biased if there are negative spillover effects from conflicts in Southern parts of the country to the control provinces. Moreover, this estimation strategy will not account for reverse causality concerns. 
Another estimation strategy is to leverage the synthetic control method as in @Abadie_2003. This approach might allow to relax the parallel trends assumption somewhat, by more flexibly allowing to control for time varying-changes in control units. 

\textbf{Bibliography}
